### 2020-06-01 writeup/experiments.tex
sec 4

Our theory leads to algorithmic consequences of practical interest.
First, we show that the diagnosis metric that we proposed in Section \ref{sec_similarity} can help predict positive or negative transfer using only STL performances.
Second, \todo{}.
Finally, we validate the three components of our theory on image and text classification tasks.

\begin{table}
\begin{minipage}[t]{.58\textwidth}
	\centering
  \begin{tabular}{c c c c c}
	\toprule
		\multirow{2}{*}{{\bf Threshold}}  & \multicolumn{2}{c}{{\bf Text
		classification}} & \multicolumn{2}{c}{{\bf ChestX-ray14}} \\
		& Precision &  Recall & Precision &  Recall \\
		\cmidrule(lr){1-1} \cmidrule(lr){2-3} \cmidrule(lr){4-5}
		0.0 & 0.596 & 1.000 & 0.593 & 1.000 \\
		0.1 & 0.756 & 0.388 & 0.738 & 0.462 \\
		0.2 & 0.919 & 0.065 & 0.875 & 0.044 \\
		% 0.3 & 1.000 & 0.004 &     - &     - \\
	\bottomrule
	\end{tabular}
	\vspace{0.1in}
	\captionof{table}{Ablation study on when should use MTL via different source/target task accuracy.}
	\label{tab:mtl_better_than_stl}
\end{minipage}
\quad
% \vspace{}
\begin{minipage}[t]{.40\textwidth}
	\centering
	\begin{tabular}{c c c}
		\toprule
		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}
		MR, SST, SUBJ, CR, MPQA, TREC\end{center}\end{minipage}} \\
		\cmidrule(lr){2-3}
		& {\bf Standard} & {\bf Alignment} \\
		\midrule
		{\bf MLP}  & > 100\% & 39\% \\
		{\bf LSTM} & 36\% & 36\% \\
		\bottomrule
		\end{tabular}
	\vspace{0.1in}
	\captionof{table}{Taskonomy experiment.}
	\label{tab:taskonomy}
\end{minipage}
\end{table}

@@

\textbf{A single-task based diagnosis metric.}
We validate the metric proposed in Section \ref{sec_similarity} for predicting positive or negative transfer based on single-task accuracies.
Table \ref{tab:mtl_better_than_stl} shows the result on both the sentiment analysis and the ChestX-ray14 datasets.
We find that using a threshold of $0.1$, our predicts correctly predicts positive or negative transfer with $75.6\%$ accuracy among 150 task pairs with $38.8\%$ recall.
We observe similar results for 91 task pairs from the ChestX-ray14 dataset.
The result shows that STL performances are indicative for the diagnosis of MTL performances.

@@

\textbf{MTL improves labeled data efficiency.}
We measure the data efficiency ratio on the sentiment analysis tasks.
We find that by performing multi-task learning, only $39\%$ of the labeled data is needed to achieve comparable performance to single-task learning over all six tasks.
If TREC is not included, we see that only $25\%$ of the labeled data is needed.

@@

\textbf{Task similarity.} \todo{}

\textbf{Data size.}
Recall that Proposition \ref{prop_data_size} shows that increasing the data size of the source task does not always improve the performance of MTL for the target task.
In Figure \ref{fig_ab_data}, we observe that for the task pair MR and SST, increasing the data size of MR is helpful initially, but then gets worse.

\textbf{Covariate shift.}
Our result provides a fine-grained insight on the covariance alignment algorithm proposed in \cite{WZR20}.
Recall that the covariance alignment procedure in \cite{WZR20} adds an additional module between the word embedding representation and the shared module.
When the source task data size is particularly large compared to the target task, then we expect that applying the covariance alignment algorithm results in a more significant gain.
We validate that the performance of aligning task covariances depends on the size of source task data.
In Figure \ref{fig_ab_cov}, we observe that the benefit from aligning task covariances becomes more significant as we increase the size of the source task dataset.

\begin{figure}[!t]
	\centering
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/task_sim_norm_lstm.pdf}
		\caption{Task model}
		\label{fig_ab_sim}
	\end{subfigure}%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_norm_sst_trec_lstm.pdf}
		\caption{Data size}
		\label{fig_ab_data}
	\end{subfigure}
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_alignment_norm_diff_all.pdf}
		\caption{Covariate shift}
		\label{fig_ab_cov}
	\end{subfigure}
	\caption{Validating the three takeaways of Section \ref{sec_insight} on sentiment analysis tasks. (a) Adding a semantically similar source task in MTL performs better than adding a dissimilar task.
	(b) As source/target data ratio increases, we observe a transition from positive to negative transfer.
	(c) As source/target data ratio increases, the performance gains from the covariance alignment procedure \cite{WZR20} increase for LSTM and MLP.}
	\label{fig_ablation}
\end{figure}

>>>>

We connect our theory to pratical problems of interest.
First, we validate the single-task based metric on text and image classification tasks.
We show that for both datasets, single-task learning results can help predict positive or negative transfer in multi-task learning.
Second, \todo{}.
Third, we measure the data efficiency ratio of multi-task learning on six sentiment analysis tasks.
Finally, we validate the three components of our theory in Section \ref{sec_insight} on text classification tasks.

\begin{table}
\begin{minipage}[t]{.58\textwidth}
	\centering
  \begin{tabular}{c c c c c}
	\toprule
		\multirow{2}{*}{{\bf Threshold}}  & \multicolumn{2}{c}{{\bf Text
		classification}} & \multicolumn{2}{c}{{\bf ChestX-ray14}} \\
		& Precision &  Recall & Precision &  Recall \\
		\cmidrule(lr){1-1} \cmidrule(lr){2-3} \cmidrule(lr){4-5}
		0.0 & 0.596 & 1.000 & 0.593 & 1.000 \\
		0.1 & 0.756 & 0.388 & 0.738 & 0.462 \\
		0.2 & 0.919 & 0.065 & 0.875 & 0.044 \\
		% 0.3 & 1.000 & 0.004 &     - &     - \\
	\bottomrule
	\end{tabular}
	\vspace{0.1in}
	\captionof{table}{Single-task learning results can help predict postive or negative transfer in multi-task learning.}
	\label{tab:mtl_better_than_stl}
\end{minipage}
\quad
% \vspace{}
\begin{minipage}[t]{.40\textwidth}
	\centering
	\begin{tabular}{c c c}
		\toprule
		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}
		MR, SST, SUBJ, CR, MPQA, TREC\end{center}\end{minipage}} \\
		\cmidrule(lr){2-3}
		& {\bf Standard} & {\bf Alignment} \\
		\midrule
		{\bf MLP}  & > 100\% & 39\% \\
		{\bf LSTM} & 36\% & 36\% \\
		\bottomrule
		\end{tabular}
	\vspace{0.1in}
	\captionof{table}{Measuring the data efficiency ratio of muli-task learning.}
	\label{tab:taskonomy}
\end{minipage}
\end{table}

@@

\textbf{Understanding MTL results via STL results.}
We show that STL results can be used to help understand MTL results
We validate the single-task based metric proposed in Section \ref{sec_similarity} for predicting positive or negative transfer in MTL.
Table \ref{tab:mtl_better_than_stl} shows the result on both the sentiment analysis and the ChestX-ray14 datasets.
We find that using a threshold of $\tau = 0.1$, the STL results correctly predict positive or negative transfer with $75.6\%$ accuracy and $38.8\%$ recall among 150 task pairs.
We observe similar results for 91 task pairs from the ChestX-ray14 dataset.
The results show that STL performances are indicative of MTL performances.

@@

\textbf{MTL improves labeled data efficiency.}
We measure the data efficiency ratio on the sentiment analysis tasks.
In Table \ref{tab:taskonomy}, we find that by performing multi-task learning, only $39\%$ of the labeled data is needed to achieve comparable performance to single-task learning over all six tasks on LSTM.
We further observe that adding the alignment module helps reduce the data efficiency ratio on MLP.
%If TREC is not included, we see that only $25\%$ of the labeled data is needed.

@@

\textbf{Task similarity.} We validate that MTL performs better when the source task is more similar to the target task.
We show the result on the sentiment analysis tasks.
For a target task, we manually select a similar task and a dissimilar task based on prior knowledge.
Figure \ref{fig_ab_sim} confirms the result.

\textbf{Data size.}
Recall that Section \ref{sec_data_size} shows that increasing the data size of the source task does not always improve the performance of MTL for the target task.
In Figure \ref{fig_ab_data}, we show that for source task MR and target task SST, there is a transition from positive to negative transfer as we increase the data size of the source task.

\textbf{Covariate shift.}
%Our result provides a fine-grained insight on the covariance alignment algorithm proposed in \cite{WZR20}.
%Recall that the covariance alignment procedure in \cite{WZR20} adds an additional module between the word embedding representation and the shared module.
When the source task data size is particularly large compared to the target task, we show that applying the covariance alignment algorithm results in more significant gains.
In Figure \ref{fig_ab_cov}, we observe that the benefit from aligning task covariances becomes more significant for LSTM and MLP as we increase the number of datapoints of the source task.

\begin{figure}[!t]
	\centering
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/task_sim_norm_lstm.pdf}
		\caption{Task similarity}
		\label{fig_ab_sim}
	\end{subfigure}%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_norm_sst_trec_lstm.pdf}
		\caption{Data size}
		\label{fig_ab_data}
	\end{subfigure}
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_alignment_norm_diff_all.pdf}
		\caption{Covariate shift}
		\label{fig_ab_cov}
	\end{subfigure}
	\caption{Validating the three takeaways of Section \ref{sec_insight} on sentiment analysis tasks. (a) Adding a semantically similar source task in MTL performs better than adding a dissimilar task.
	(b) As source/target data ratio increases, we observe a transition from positive to negative transfer.
	(c) As source/target data ratio increases, the performance gains from the covariance alignment procedure \cite{WZR20} increase for LSTM and MLP.
	The $y$-axis measures the performance of MTL minus STL.}
	\label{fig_ablation}
\end{figure}

### 2020-06-01 writeup/transfer.tex
sec 4.3
sec 4

Based on Lemma \ref{lem_cov_shift_informal}, we rigorously explain the phenomena of negative transfer in multi-task learning.
We explain from three perspectives, including \textit{task model}, \textit{data size} and \textit{covariate shift}.
We show how negative transfer occurs as a result in a simplified model and describe the algorithmic consequence.

\subsection{Task Model}\label{sec_similarity}

@@

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Task model}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Data size}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{Three takeaways of our theory in Section \ref{sec_insight}.
	(a) As task model distance increases, we observe a transition from positive to negative transfer  (Proposition \ref{prop_dist_transition}).
	(b) As source/target data ratio increases, we observe a transition from positive to negative transfer (Proposition \ref{prop_data_size}).
	(c) As covariate shift gets more severe (measured by increasing $\lambda$), MTL performance gets worse (Proposition \ref{prop_covariate}).}
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

\subsection{Data Size}
In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the data sizes are equal for all tasks \cite{B00,M06,MPR16}.
More generally, such results are still applicable when all task data are being added simultaneously.
On the other hand, for many applications of multi-task learning, the data sources are usually heterogeneous.
For such settings, we have observed that adding more labeled data from one task does not always help.
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using the isotropic model, we consider what happens if we vary the source task data size.
Figure \ref{fig_size} provides a simulation result for such a setting.
We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
Our result below explains this phenomenon.

>>>>

Based on Lemma \ref{lem_cov_shift_informal}, we rigorously explain the phenomena of negative transfer in multi-task learning.
We explain from three perspectives, including \textit{task similarity}, \textit{data size} and \textit{covariate shift}.
We show how negative transfer occurs as a result in a simplified model and describe the algorithmic consequence.

\subsection{Task Similarity}\label{sec_similarity}

@@

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Task similarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Data size}
		\label{fig_size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{Three takeaways of our theory in Section \ref{sec_insight}.
	(a) As task model distance increases, we observe a transition from positive to negative transfer  (Proposition \ref{prop_dist_transition}).
	(b) As source/target data ratio increases, we observe a transition from positive to negative transfer (Proposition \ref{prop_data_size}).
	(c) As covariate shift gets more severe (measured by increasing $\lambda$), MTL performance gets worse (Proposition \ref{prop_covariate}).
	The $y$-axis measures the performance of MTL minus STL.}
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

\subsection{Data Size}\label{sec_data_size}

In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the data sizes are equal for all tasks \cite{B00,M06,MPR16}.
More generally, such results are still applicable when all task data are being added simultaneously.
On the other hand, for many applications of multi-task learning, the data sources are usually heterogeneous.
For such settings, we have observed that adding more labeled data from one task does not always help.
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using the isotropic model, we consider what happens if we vary the source task data size.
Figure \ref{fig_size} provides a simulation result for such a setting.
We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
Our result below explains this phenomenon.
